|

### 2021-09-13 03-Chp-Systematic-Review.Rmd
End of day:

@@

@@

<!-- Taken from another paper: "A potential non-linear association was evaluated using a restricted cubic spline model with three knots at the 10th, 50th and 90th percentile of frequency of the exposure [@orsini2012meta]" -->
<!-- TODO Do I initially want to fit a linear relationship across categories and  -->
A <!-- TODO confirm type of spline --> spline model was fitted to allow for a non-linear relationship, for example a U or J-shaped relationship, where low and high levels of an exposure can have an effect versus the "normal" reference dose. Reference doses were defined _a priori_ as the cut-off of the "Normal"/"Optimal" categories for each fractions, as detailed in Table \@ref(tab:lipidLevels-table). Under this approach, the reference dose was defined as 200 mg/dL for total cholesterol, 100 mg/dL for LDL-c, 40 mg/dL for HDL-c, and 150 mg/dL for triglycerides.

@@

Further, given the importance of visualising the potential biases of a result alongside the result itself, a new visualisation tool was designed to allow for "paired" forest plots (as recommended by the ROB2 publication). <!-- TODO CITATION NEEDED --> This tool was developed as an adjunct to this thesis to aid in creating standard risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other reserachers to design a modular method for creating these plots.[@zotero-14999] <!-- Apparently collaboration is a big thing they want to see --> A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this analysis were created using this tool.

@@

Preprints are a valuable evidence source (see Introduction, Section \@ref()) but their inclusion in a systematic reivew. As part of a study within this larger review, I also to explore the additional evidental value of including preprints in the meta-analysis, assessed using the fixed effect weight from a standard meta-analysis.

@@

@@

@@

@@

\blandscape{}
<!----------------------------------------------------------------------------->
(ref:studyCharacteristics-caption) studyCharacteristics

(ref:studyCharacteristics-scaption) studyCharacteristics

@@

Many of the included studies were at high-risk of bias.

The one randomised controllled trial was deemed to be a low risk of bias,

@@

### Meta-analysis of statin use and hypercholesteroleamia

@@

Statin use was associated with a reduce risk of all-cause dementia (`r estimate(0.76,0.66,0.88)`, Figure \@ref(fig:fpStatins)).

&nbsp;
<!----------------------------------------------------------------------------->
```{r primaryFigures, include = FALSE}
```

(ref:fpStatins-cap) __Forest plot showing effect of statins:__

(ref:fpStatins-scap) Forest plot showing effect of statins

```{r fpStatins, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:fpStatins-cap)', out.width='100%', fig.scap='(ref:fpStatins-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->
&nbsp;

@@

Several studies were excluded from the dose-response meta-analysis, as the number of cases/controls per dose group could not be calculated and the corresponding author for the study did not respond to clarification requests. The results from the dose response analysis can be seen in Figure \@ref(fig:dose-response).

<!-- TODO Include image -->

@@

@@

As show in Figure \@ref(prisma-flow-fig), the number of results returned by the preprint searching, was not substantial (bioRxiv = 256, medRxiv = 0). <!-- TODO check -->

@@

### Risk of bias due to missing evidence

@@

some evidence that age has impact on observed lipid-dementia relationship - use to link

@@

@@

@@

@@

Firstly, as this study design is relatively new, particularly when compared to randomised trials or cohort studies, the process and tools for systematically assessing them are not as well developed. A key example of this is in the absence of validated search filters for Mendelian randomisations studies. This limitation is further complicated by the varying terminology used to describe the method, particularly in the early years of it's application.

>>>>

#### Combining across groups

Following best practice, where summary data was presented across two groups (e.g. age at baseline stratified by hypercholesterolemai status), the following approach was used to combine the groups <!-- TODO Cite cochrane handbook here -->

<!-- TODO Insert equations here -->

\begin{equation}
N = N_1 + N_2
  (\#eq:combiningGroups1)
\end{equation}

\begin{equation}
Mean = \frac{(N_1M_1 + N_2N_2)}{(N_1 + N_2)}
  (\#eq:combiningGroups2)
\end{equation}

\begin{equation}
SD = \sqrt{\frac{(N_1-1)SD_1^2 + (N_2-1)SD_2^2 + \frac{N_1N_2}{N_1 + N_2}(M_1^2 + M_2^2 - 2M_1M_2)}{N1 + N2 -1}}
  (\#eq:combiningGroups3)
\end{equation}

This was implemented in a systematic manner, with the raw group data being extracted and a cleaning script written to combine the groups for analysis. where more than one group

@@

<!-- TODO Check these against those in Ben's thesis paper -->

@@

<!-- Taken from another paper: "A potential non-linear association was evaluated using a restricted cubic spline model with three knots at the 10th, 50th and 90th percentile of frequency of the exposure [@orsini2012meta]" -->
<!-- TODO Do I initially want to fit a linear relationship across categories and  -->
A <!-- TODO confirm type of spline --> spline model was fitted to allow for a non-linear relationship, for example a U or J-shaped relationship, where low and high levels of an exposure can have an effect versus the "normal" reference dose. Reference doses were defined _a priori_ as the cut-off of the "Normal"/"Optimal" categories for each fractions, as detailed in Table \@ref(tab:lipidLevels-table). Under this approach, the reference dose was defined as 200 mg/dL for total cholesterol, 100 mg/dL for LDL-c, 40 mg/dL for HDL-c, and 150 mg/dL for triglycerides. <!-- TODO Each bracket should have an upper and lower end -->

@@

Further, given the importance of visualising the potential biases of a result alongside the result itself, a new visualisation tool was designed to allow for "paired" forest plots (as recommended by the ROB2 publication).[@sterne2019] This tool was developed as an adjunct to this thesis to aid in creating standard risk-of-bias figures,[@mcguinness2020robvisPaper] and the "paired" forest plot functionality grew out of a collaboration with other reserachers to design a modular method for creating these plots.[@zotero-14999] <!-- Apparently collaboration is a big thing they want to see --> A summary of this tool is contained in Appendix \@ref(appendix-robvis), and all forest plots presented in this analysis were created using this tool.

@@

Preprints are a valuable evidence source (see Introduction, Section \@ref(diverse-sources-preprints)) but their inclusion in a systematic review.

As part of a study within this larger review, I also to explore the additional evidental value of including preprints in the meta-analysis, assessed using the fixed effect weight from a standard meta-analysis.

@@

<!-- TODO Compare findings for two papers from Rantanen -->

@@

<!-- TODO Finish PRISMA -->

@@

Of note, several studies provided evidence on a lipid fraction as part of a wider study on Mets. Raffatin/Ng

@@

\blandscape{}
<!----------------------------------------------------------------------------->
(ref:studyCharacteristics-caption) Characteristics of included studies, stratified by study design.

(ref:studyCharacteristics-scaption) Characteristics of included studies

@@

Many of the included results were assessed to be at high risk of bias.

The one randomised controlled trial was deemed to be a low risk of bias,

Several synthesis were also assessed to be at risk of bias due to missing evidence.

@@

### Statins

```{r primaryFigures, include = FALSE, message=FALSE,warning=FALSE}
```

@@

In the two included RCTs, statin use was not associated with dementia outcomes

<!----------------------------------------------------------------------------->
(ref:statinsRCT-cap) __Random effects meta-analysis of statins on all-cause dementia:__

(ref:statinsRCT-scap) Random effects meta-analysis of statins on all-cause dementia

```{r statinsRCT, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:statinsRCT-cap)', out.width='100%', fig.scap='(ref:statinsRCT-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forest_statins_rct.png"))
```
<!----------------------------------------------------------------------------->

In NRSI, statin use was associated with a reduce risk of all-cause dementia (`r estimate(0.76,0.66,0.88)`, Figure \@ref(fig:statinsNRSI)).

(ref:statinsNRSI-cap) __Forest plot showing effect of statins:__

(ref:statinsNRSI-scap) Forest plot showing effect of statins

```{r statinsNRSI, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:statinsNRSI-cap)', out.width='100%', fig.scap='(ref:statinsNRSI-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

In Mendelian randomisation studies, a similar blah blah blah

(ref:statinsMR-cap) __Forest plot showing effect of statins:__

(ref:statinsMR-scap) Forest plot showing effect of statins

```{r statinsMR, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:statinsMR-cap)', out.width='100%', fig.scap='(ref:statinsMR-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

### Hypercholesterolemia and lipid levels

(ref:hypercholNRSE-cap) __Forest plot showing effect of statins:__

(ref:hypercholNRSE-scap) Forest plot showing effect of statins

```{r hypercholNRSE, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:hypercholNRSE-cap)', out.width='100%', fig.scap='(ref:hypercholNRSE-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

Several studies presented results per SD change in a lipid fraction.

(ref:lipidsNRSE-cap) __Forest plot showing effect of statins:__

(ref:lipidsNRSE-scap) Forest plot showing effect of statins

```{r lipidsNRSE, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:lipidsNRSE-cap)', out.width='100%', fig.scap='(ref:lipidsNRSE-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

@@

Several studies were excluded from the dose-response meta-analysis, as the number of cases/controls per dose group could not be calculated and the corresponding author for the study did not respond to clarification requests. The results from the dose response analysis can be seen in Figure \@ref(fig:lipidsDoseResponse).

(ref:lipidsDoseResponse-cap) __Forest plot showing effect of statins:__

(ref:lipidsDoseResponse-scap) Forest plot showing effect of statins

```{r lipidsDoseResponse, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:lipidsDoseResponse-cap)', out.width='100%', fig.scap='(ref:lipidsDoseResponse-scap)'}
knitr::include_graphics(file.path("figures/sys-rev/forester_statins_any.png"))
```
<!----------------------------------------------------------------------------->

@@

Planned to investigate effect of sex and age at baseline.

Plus other such as education level and baseline cognitive scores, but data were either not reported for most studies or when reported, were too diverse to synthesize.

@@

I had intended to assess the added evidental value of including preprints as a literautre source.

As show in Figure \@ref(prisma-flow-fig), the number of results returned by the preprint searching was not substantial (bioRxiv = 256, medRxiv = 0). <!-- TODO check numbers -->

No unique preprints were identified as part of this review, with the published version of the few/sole <!-- TODO check which --> relevant preprint(s) being captured by the main search.

@@

@@

There was some evidence of protective effect of statins on all-cause and Alzheimer's disease dementia when looking at solely at observational studies. This finding was not supported by evidence from the single available RCT, or by studies that emulated statin treatment using a genetic proxy, suggesting that these findings are a

In addition,  there were some concerns over the potential for estimates to be missing from the meta-analysis of observational studies not at random, given the preferential reporting of

Some evidence that age has impact on observed lipid-dementia relationship - use to link

@@

<!-- The findings on high TC in midlife being associated with increased risk of late-life AD are still limited to the few studies that report relevant data from Scandinavia and the United States. The wider literature shows that level of TC in populations varies according to diet, urbanization, ethnicity, and income [56, 57]. from Anstey 2017 -->

Questions over missing results - evidence from one of the conference abstract analysis pairs that a non-significant results are being suppressed.[@yamada2009; @yamada2009a] Tie this in with

@@

<!-- TODO Discuss information here that I responded to Julian with re: % of preprints that published -->

@@

Additionally many of the studies suggested a link between, but then in the sensitivity analyses or in the Discussion disclaim that when adjusted for ApoE4, any association was attenuated to the null. This point is important given one of the core assumptions of

<!-- TODO Diagram showing three assumptions of MR, comparing it to a graph showing the invalidation of no pleiotopy assumption -->

@@

Firstly, this study design is relatively new, particularly when compared to randomised trials or cohort studies. Figure \@ref(fig:typeByYear) demonstrates that MR studies only begin to appear in the evidence base much later than NRSE/NRSI. As such, the process and tools for systematically assessing them are not as well developed, likely due to the limited availabilt of large scale GWAS datasets needed for two sample MR. A key example of this is in the absence of validated search filters for Mendelian randomisations studies. This limitation is further complicated by the varying terminology used to describe the method, particularly in the early years of it's application.

### 2021-09-13 04-Chp-CPRD-Analysis.Rmd
End of day:

No evidence for an effect of lipid regulating agents effect on the risk of Alzheimer's disease was found, with the exception of a slightly increased risk in those prescribed a certain type of drugs called fibrates. In contrast, I found an increased risk of vascular and other (i.e. non-Alzheimer's) dementia with lipid regulating agent use.

@@

As part of a Cox proportional hazard model, there is the option to use either absolute time in cohort or participants age as the time scale of interest.[@lamarca1998; @gail2009; @pencina2007] A model using age as the time axis inherently accounts, or adjusts, for participants age as a potential confounded of the exposure-outcome relationship. The main analyses presented all used age as the time axis.

@@

Due to the wealth of data available on statins as a lipid regulating agent, control outcomes were chosen in reference to this drug class. Muscular backpain was chosen as a negative control outcome in this analysis. Despite observational analyses suggesting a link between statins and muscle pain (as opposed to more serious complications such as myopathy),[@selva-ocallaghan2018] systematic reviews of the adverse events of statin use[@collins2016a] and N-of-1 trials explicitly exploring the effect of statin use on muscle pain[@herrett2021] have found no evidence of an effect. As such, if no evidence of an effect of statins on backpain is found in the analysis, then confidence in the results for the dementia outcomes is increased.

@@

To explore whether any observed effect for the statin analysis varied by statin property, a sensitivity analysis was performed, stratifying by statin lipophilicity.

@@

This previous study used a propensity matching approach to estimate the effect of statins on a range of outcomes, and a protective effect of statin use on all-cause dementia (`r estimate(0.81, 0.69,0.96,type = "HR")`) and non-AD dementia (`r estimate(0.82,0.69, 0.97,type = "HR")`), but no evidence of an effect on AD (`r estimate(0.81,0.49, 1.35, type="HR")`).

@@

No evidence was found for an effect of lipid-regulating agents on probable (`r probad_text`) and possible (`r possad_text`) Alzheimer's disease when compared with no treatment, with the sole exception of fibrates on probable Alzheimer's disease (`r probad_fib_text`).

@@

When stratifying by statin properties, hydrophilic statins were less harmful in the any, vascular and other dementia outcomes compared to lipophilic statins (Figure \@ref(fig:statinTypeFig)). Additionally, in the AD outcomes, hydrophilic statins had a small protective effect, compared to the absence of evidence for an effect for lipophilic statins.

@@

Lipid-regulating agents showed no evidence of an effect on probable and possible Alzheimer's disease when compared with no treatment, but were associated with increased risk of an all-cause dementia, vascular dementia and other dementia diagnosis. The effect observed in each case was driven by the statin subgroup, which included a substantial majority of participants. For the other drug classes, no association was found with any outcome, with two exceptions being that ezetimibe was associated with increased risk of vascular and other dementia, while fibrates were associated with increase risk of all-cause dementia and probable Alzheimer's disease.

@@

It found no evidence for an effect of lipid-regulating agents on probable or possible Alzheimer’s disease. However, lipid regulating agent use was associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations were driven by those observed in the statin subgroup, which comprised the majority of exposed participants in this cohort.

>>>>

Little evidence for an effect of lipid regulating agents effect on the risk of Alzheimer's disease was found, with the exception of a slightly increased risk in those prescribed a certain type of drugs called fibrates. In contrast, I found an increased risk of vascular and other (i.e. non-Alzheimer's) dementia with lipid regulating agent use.

@@

As part of a Cox proportional hazard model, there is the option to use either absolute time in cohort or participants age as the time scale of interest.[@lamarca1998; @gail2009; @pencina2007] A model using age as the time axis inherently accounts, or adjusts, for participants age as a potential confounder of the exposure-outcome relationship. As such, the main analyses presented all used age as the time axis.

@@

Due to the wealth of data available on statins as a lipid regulating agent, control outcomes were chosen in reference to this drug class. Muscular backpain was chosen as a negative control outcome in this analysis. Despite observational analyses suggesting a link between statins and muscle pain (as opposed to more serious complications such as myopathy),[@selva-ocallaghan2018] systematic reviews of the adverse events of statin use[@collins2016a] and N-of-1 trials explicitly exploring the association of statin use with muscle pain[@herrett2021] have found little evidence supporting an effect. As such, if statins are not associated with backpain in this analysis, then confidence in the results for the dementia outcomes is increased.

@@

To explore whether any observed associations in the statin analysis varied by statin property, a sensitivity analysis was performed, stratifying by statin lipophilicity.

@@

This previous study used a propensity matching approach to estimate the association of statins with a range of outcomes, and found evidence for a protective effect of statin use on all-cause dementia (`r estimate(0.81, 0.69,0.96,type = "HR")`) and non-AD dementia (`r estimate(0.82,0.69, 0.97,type = "HR")`), but little evidence of an effect on AD (`r estimate(0.81,0.49, 1.35, type="HR")`).

@@

My results show litte evidence was found for an effect of lipid-regulating agents on probable (`r probad_text`) and possible (`r possad_text`) Alzheimer's disease when compared with no treatment, with the sole exception of fibrates on probable Alzheimer's disease (`r probad_fib_text`).

@@

When stratifying by statin properties, hydrophilic statins were less harmful in the any, vascular and other dementia outcomes compared to lipophilic statins (Figure \@ref(fig:statinTypeFig)). Additionally, in the AD outcomes, hydrophilic statins were associated with a small reduction in risk, compared to the weak evidence for an effect for lipophilic statins.

@@

Lipid-regulating agents showed little evidence of an effect on probable and possible Alzheimer's disease when compared with no treatment, but were associated with increased risk of an all-cause dementia, vascular dementia and other dementia diagnosis. The effect observed in each case was driven by the statin subgroup, which included a substantial majority of participants. For the other drug classes, no association was found with any outcome, with two exceptions being that ezetimibe was associated with increased risk of vascular and other dementia, while fibrates were associated with increase risk of all-cause dementia and probable Alzheimer's disease.

@@

It found little evidence for an effect of lipid-regulating agents on probable or possible Alzheimer’s disease. However, lipid regulating agent use was associated with an increased risk of all-cause, vascular and other dementias. In all cases, the estimated associations were driven by those observed in the statin subgroup, which comprised the majority of exposed participants in this cohort.

### 2021-09-13 05-Chp-IPD-Analysis.Rmd
End of day:

@@

Talk about how different covariates were managed across cohorts,

<!-- TODO Read this[@riley2020] -->

@@

* Description (including observation period, numbers, numbers with outcome, etc)
* Whether they are a known genetically at-risk cohort
*

@@

@@

@@

#### Memento

@@

@@

@@

@@

Across all cohorts, data cle

@@

For heterogeneity across the cohorts, the total time at risk was investigated.

@@

@@

@@

This was done to

CLustering within studies was accounted for, given the evidence that ...[@abo-zaid2013]

@@

@@

@@

@@

@@

While this is not unexpected, given that a review of IPD studies published between 1987 and 2015 found that fewer than half managed to obtained data from greater than 80% of studies, and that in many cases, the exact percentage of studies for which data was obtained was not accurately reported.

There are likely several reasons for this.

@@

In future, it could have been worth running a "study within a review", such as that used to examine the best method [@godolphin2019a]

>>>>

Justify analysis based on findings of systematic review - suggestion of variation of effect by age, specifically

@@

Talk about how different covariates were managed across cohorts, 
c
<!-- TODO Read this[@riley2020] -->

@@

* Description (including observation period, numbers, numbers with outcome, etc)
*

@@

Has only 8 events

@@

Participants were included from the first date of lipid measurement, so no issues with immortal time bias as discussed in Chapter ...

Additionally, ther number of participants is larger in this analysis as there is no restriction on the level which lipids should be in order to be included in the analysis.

the

@@

@@

Likley had dementia at baseline

#### Memento
Excluded as criteria for entry was outpatients from memory clinic (unlikely to be dementia free at baseline)

Cross-sectional data only

NICOLA study - only had cross-sectional data

Data provided but only had ever high cholesterol as a binary variable. Not compatible with

@@

LBC - 1936

Taiwan etc,

@@

<!-- TODO Will need some method to define how/from where the different cohorts were obtained. -->

@@

Across all cohorts, data cleaning was performed in a similary manner, using commonly named variables, so that a single model could be applied using functional programming.

The one exception to this is the CPRD data, which was held in a different system to the rest.

The advantage of this approach is that it reduces the liklihood of errors in model mis-speciication if needing to change variables names from cohort to cohort.

@@

For heterogeneity across the cohorts, the total time-at-risk was investigated.

@@

There is some concerns about performing risk of bias assessments on your own analysis, and so

@@

<!-- TODO talk about different methods - see here: https://onlinelibrary.wiley.com/doi/10.1002/sim.8516 -->

@@

This was done to aid interpretability of the outcome.

Similary,

Clustering within studies was accounted for, given the evidence that ...[@abo-zaid2013]

@@

Hazard ratios were utilised to

In order to investigate the interaction of sex with

A two-stage model was used out of necessity, with the one-stage model being recluded by the different datasets being in protected data silos.

A discrete proportional hazards model was employed to account for the interval censoring introduced by design o fthe longitudinal cohort studies.[@wang2017]

<!-- TODO Figure showing the outline of interval censored studies. -->

@@

Data

@@

Useful citation for discussion[@levis2021]

@@

<!-- TODO Plot showing when datasets requested were published and comparing this with the -->

<!-- TODO text -->

@@

While this is not unexpected, given that a review of IPD studies published between 1987 and 2015 found that fewer than half managed to obtained data from greater than 80% of studies, and that in many cases, the exact percentage of studies for which data was obtained was not accurately reported.
test
There are likely several reasons for this.

@@

In future, it could have been worth running a "study within a review", such as that used to examine the best method [@godolphin2019a] to identify and access study data.

Previous SWARs have
